---
title: "Creating Data in R"
comment: "*introduction to data generation and simulation*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , echo = FALSE, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      error = TRUE)
library(tidyverse)
library(skimr)
```


# Objectives

This lesson outlines some examples for creating data in R. Being able to generate your own datasets in R makes it easier to test code and create reproducible examples. We're going to cover the following packages for creating data in R: 

1. [fakeR](https://cran.r-project.org/web/packages/fakeR/fakeR.pdf)  
2. [fabricatr](https://declaredesign.org/r/fabricatr/articles/getting_started.html)  
3. [charlatan](https://docs.ropensci.org/charlatan/)
4. [fakir](https://thinkr-open.github.io/fakir/)

Read more about data simulation [here](https://rviews.rstudio.com/2019/07/01/imagine-your-data-before-you-collect-it/) and [here](https://rviews.rstudio.com/2020/09/09/fake-data-with-r/) on the RStudio blog.

# Load packages

Install and load packages below:

```{r packages, message=FALSE, warning=FALSE}
install.packages(c("fakeR", "fabricatr",
                   "charlatan", "fakir",
                   "tidyverse"))
library(fakeR)
library(fabricatr)
library(charlatan)
library(fakir)
library(tidyverse)
```



# Materials (coming soon!)

The slides for this presentation are [here]().

The RStudio.Cloud project [here]().


# `fakeR`

The `fakeR` package has been.

```{r}

```


# `fabricatr`

# `charlatan`

# `fakir`


