---
title: "Scraping Wikipedia Tables"
comment: "*advanced wrangling example*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , include=FALSE}
library(tidyverse)
library(skimr)
# chunk options ----
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      tidy = FALSE,
                      fig.path = "img/")
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
```

# Objectives

This document outlines some advanced data wrangling with [`dplyr`](https://dplyr.tidyverse.org/index.html), [`tidyr`](https://tidyr.tidyverse.org/index.html), and [`stringr`](https://stringr.tidyverse.org/index.html). Read more in the [Data Transformation](https://r4ds.had.co.nz/transform.html) chapter of R for Data Science.

## Load packages

These are all part of the `tidyverse`, so we install and load this meta-package below.

```{r packages}
# install.packages("tidyverse")
library(tidyverse)
```

# Pharmaceutical Companies

In the slides we had well-formatted dataset. In these exercises, we're going to import some data from the wild (Wikipedia), because most data aren't ready to visualize and model when we first get them.

## Import the data {.tabset}

I Googled "largest pharmaceutical companies" and found [this wikipedia page](https://en.wikipedia.org/wiki/List_of_largest_biotechnology_and_pharmaceutical_companies). 

These packages will help us scrape the data in the table and manipulate it in R. 

```{r scrape-packages, message=FALSE, warning=FALSE}
library(xml2)
library(rvest)
```


### Read HTML

The first function I'll use comes from the `xml2` [package](https://blog.rstudio.com/2015/04/21/xml2/). `xml2::read_html()` loads the html from the wikipedia page into an R object I call `wiki_html`.

```{r wiki_html}
wiki_html <- xml2::read_html("https://en.wikipedia.org/wiki/List_of_largest_biotechnology_and_pharmaceutical_companies")
```

We should check the structure of new objects, so we know what we're working with.

```{r structure_wiki_html}
wiki_html %>% str()
```

I can see this is a list of two objects (a `node` and a `doc`).

### HTML Nodes (css)

I want the html `node`, so I will use the `html_nodes()` function from the `rvest` [package](https://github.com/hadley/rvest). 

The `css` argument is set to `"table"`. Once again I check the structure of the output object.

```{r wiki_html_tables}
wiki_html_tables <- wiki_html %>% rvest::html_nodes(css = "table")
wiki_html_tables %>% str()
```

This is a list of three lists, each of them `xml_node`s. 

In order to find the relevant tables in the `wiki_html_tables` object, I need to be able to search on something. Fortunately, the `base::grep()` function can be used in combination with sub-setting to extract the `relevant_tables` from `wiki_html_tables`.

Get the relevant tables from the `xml_nodeset` in `wiki_html_tables`.

```{r relevant_tables}
relevant_tables <- wiki_html_tables[base::grep("rank", wiki_html_tables)]
relevant_tables %>% str()
```

This returned yet another list of lists! (another set of `node` and `doc`). 

### Extract HTML table

Now I need to use `rvest::html_table()` function (combined with some bracket sub-setting) to explore this object and learn about it's contents (read more about how this works [here](https://dcl-wrangle.stanford.edu/rvest.html#scrape-data-with-rvest)). 

I will start with position `[[1]]` and set `fill = TRUE`. 

```{r test-html_table}
rvest::html_table(relevant_tables[[1]], 
                  fill = TRUE) %>% utils::head()
```

This looks like the table I want! Let's assign it to a data frame and get wrangling!

```{r TopPharmCompRaw}
TopPharmCompRaw <- rvest::html_table(relevant_tables[[1]], 
                  fill = TRUE)
```

```{r export-TopPharmCompRaw, echo=FALSE, error=TRUE, message=FALSE, warning=FALSE}
library(readr)
# little fail-safe here to export the current data as-is and read it back in just in case the data change in the future... 
# write_csv(x = TopPharmCompRaw, path = "data/2020-11-24-TopPharmCompRaw.csv")
# read back in
TopPharmCompRaw <- read_csv(file = "https://bit.ly/3orq1YT")
```


## Outline before you start 

We have a raw dataset now, and we should make a 'to-do' list of what we want the data to look like. I start with data tidying, then move into changes for the individual variables. 

# Wrangle Data 

These data are in a wide format, with each `market_cap_in_` spread across columns. 

```{r view-TopPharmCompRaw}
TopPharmCompRaw
```


## Pivot data {.tabset}

Ideally, we want two variables, `year` and `market_cap_us_bil`. 

### `pivot_longer`

We can re-shape `TopPharmComp` with `tidyr::pivot_longer()`, but first we need all the variables in the *same* format. We can do this with `mutate_if()`.

```{r check-pivot_longer}
TopPharmCompRaw %>% 
  mutate_if(is.numeric, as.character) %>% 
  tidyr::pivot_longer(cols = starts_with("Market Cap in"), 
                      names_to = "year", 
                      values_to = "market_cap_us_bil")
```

This looks correct--let's assign it to `TidyTopPharmComp` (because it's a new data structure).

```{r assign-TidyTopPharmComp}
TidyTopPharmComp <- TopPharmCompRaw %>% 
  mutate_if(is.numeric, as.character) %>% 
  tidyr::pivot_longer(cols = starts_with("Market Cap in"), 
                      names_to = "year", 
                      values_to = "market_cap_us_bil")
```

## Format Columns {.tabset}

Follow along on each tab for the steps to wrangle the column names:

1. Each variable in it's own column:  
   - i.e. type of company in it's own column (`B` – Biotechnology company, `P` – Pharmaceutical company)
   - stock exchange identifier in it's own column (i.e. `NYSE`, `NASDAQ`, etc.) 
2. Properly formatted values (numeric, factor, etc.)   
3. Missing variables formatted correctly (replace `-` with `NA`)

### String manipulations

We can use `janitor::clean_names()`, but first we should do some string manipulation to remove the extra characters from the column names. 

```{r cleaned_names}
TidyTopPharmComp %>% 
    # get a vector of 'dirty' names
    names() %>% 
    # remove citations in brackets [ ]
    stringr::str_remove_all(string = ., pattern = "\\[[0-9]\\]") %>%
    # replace (USD billions)
    stringr::str_replace_all(string = ., pattern = "\\s*\\([^\\)]+\\)", 
                             replacement = "_us_bil") %>% 
    # remove bracket from Rank
    stringr::str_remove_all(string = ., pattern = "\\[|\\]") %>% 
    # remove alpha numeric from end of Rank
    stringr::str_remove_all(string = ., pattern = "N 1$") %>% 
    # make all lowercase
    stringr::str_to_lower() -> cleaned_names
cleaned_names
```

### `janitor::clean_names`

Now we can apply the new names (`cleaned_names`) to `TidyTopPharmComp` and use `janitor::clean_names()` to standardize them. 

```{r clean_names}
TidyTopPharmComp %>% 
    set_names(nm = cleaned_names) %>% 
    janitor::clean_names()
```

### Assign to data frame

If this looks acceptable, we can assign to `TopPharmComp`.

```{r TopPharmComp}
TopPharmComp <- TidyTopPharmComp %>% 
    set_names(nm = cleaned_names) %>% 
    janitor::clean_names()
```



## Define Variables {.tabset}

We want to split up the following variables into their own columns: 

1. Year as a four-number digit `year`
2. `company_type` = `[P]` for `Pharmaceutical` or `[B]` for `Biotechnology`  
3. `stock_exch` = `NSYE` (New York Stock Exchange), `NASDAQ` (National Association of Securities Dealers Automated Quotations), `FWB` (Frankfurt Stock Exchange), `TYO` (Australian Securities Exchange), `TSX` (Toronto Stock Exchange), and `SIX` (Swiss Exchange).
4. `stock_id` = acronym for each company on `stock_exch`  
5. `largest_market_cap_date` = date from `largest_market_cap_us_bil`  
6. `company_name` a variable containing *only* the company name 
7. a `ranking` variable with a numerically coded rank

### `year`

We can also wrangle the `market_cap_year` variable so it only contains the four-number year.

```{r check-market_cap_year}
TopPharmComp %>% 
  # remove characters
  dplyr::mutate(year = str_remove_all(string = year, pattern = "\\D"),
                year = str_sub(string = year, start = 1L, end = 4L),
                # make numeric
                year = as.integer(year))
```

Let's assign `year` to `TopPharmComp`

```{r assign-TopPharmComp}
TopPharmComp %>% 
  # remove characters
  dplyr::mutate(year = str_remove_all(string = year, pattern = "\\D"),
                year = str_sub(string = year, start = 1L, end = 4L),
                # make numeric
                year = as.integer(year)) -> TopPharmComp
```

### `company_type`

Here we `separate` the `company` column into two new columns, then use the `company_name_etc` to create `company_type`

```{r check-company_type}
TopPharmComp %>% 
  # separate company into 2 columns
  separate(col = company, into = c("company_name_etc", "company_etc"), 
           sep = "\\]", 
           remove = FALSE) %>% 
  # create categories
  mutate(company_type = case_when(
    str_detect(string = company_name_etc, pattern = "\\[P$") ~ "Pharma", 
    str_detect(string = company_name_etc, pattern = "\\[B$") ~ "Biotech")) %>% 
  # check our work
  dplyr::select(contains("comp"))
```

These categories look are correct, so we can assign `company_type` to `TopPharmComp`.

```{r assign-company_type}
TopPharmComp <- TopPharmComp %>% 
  # separate company into 2 columns
  separate(col = company, into = c("company_name_etc", "company_etc"), 
           sep = "\\]", 
           remove = FALSE) %>% 
  # create categories
  mutate(company_type = case_when(
    str_detect(string = company_name_etc, pattern = "\\[P$") ~ "Pharma", 
    str_detect(string = company_name_etc, pattern = "\\[B$") ~ "Biotech")) %>% 
  # remove company_name_etc
  select(-company_name_etc)
TopPharmComp
```

### `stock_exch` & `stock_id`

We will use the `company_etc` column to create `stock_exch` and `stock_id`. We need to make sure we trim the whitespace from `stock_id` (after the )

```{r check-stock_exch-stock_id}
TopPharmComp %>% 
  # first separate the columns
    separate(col = company_etc, 
             into = c("stock_exch", "stock_id"), 
             sep = ":", remove = FALSE) %>% 
  # check our work
  select(company,
         company_etc, 
         contains("stock"))
```

The `sep` argument helps us a lot here, so we can create two new variables with minimal steps! We'll assign to `TopPharmComp`.

```{r assign-stock_exch-stock_id}
TopPharmComp %>% 
  # first separate the columns
    separate(col = company_etc, 
             into = c("stock_exch", "stock_id"), 
             sep = ":", remove = FALSE) %>% 
  select(rank, 
         contains("comp"),
         contains("stock"),
         everything(), 
         -company_etc) -> TopPharmComp
TopPharmComp
```

### `largest_market_cap_date`

Here we'll use `dplyr::rename()` to change `largest_market_cap_us_bil` to `_original` so we can check our work with `remove = FALSE`.

```{r check-largest_market_cap_date}
TopPharmComp %>% 
  # rename column to largest_market_cap_original
  dplyr::rename(largest_market_cap_original = largest_market_cap_us_bil) %>% 
  # separate into two columns
  separate(col = largest_market_cap_original, 
           into = c('largest_market_cap_us_bil', 'largest_market_cap_date'),
           sep = "\\(", 
           remove = FALSE) %>% 
  # remove parenthesis
  mutate(largest_market_cap_date = 
           str_remove_all(largest_market_cap_date, pattern = "\\)$")) %>% 
  # check our work
  select(contains("large"))
```

Now we have the month and year separated into `largest_market_cap_date`--we can assign to `TopPharmComp`

```{r assign-largest_market_cap_date}}
 TopPharmComp %>% 
  # rename column to largest_market_cap_original
  dplyr::rename(largest_market_cap_original = largest_market_cap_us_bil) %>% 
  # separate into two columns
  separate(col = largest_market_cap_original, 
           into = c('largest_market_cap_us_bil', 'largest_market_cap_date'),
           sep = "\\(", 
           remove = FALSE) %>% 
  # remove parenthesis
  mutate(largest_market_cap_date = 
           str_remove_all(largest_market_cap_date, pattern = "\\)$")) %>% 
  # reorganize
  select(rank, 
         contains("comp"),
         contains("stock"),
         contains("large"),
         everything(), 
         -largest_market_cap_original) -> TopPharmComp
TopPharmComp
```


### `company_name`

We will create a `company_name` variable using a different `sep` argument here, and we clean the `company_name` variable with `str_remove_all()`

```{r check-company_name}
TopPharmComp %>% 
  # separate company into 2 columns
  separate(col = company, into = c("company_name", "company_junk"), 
           sep = "\\[",
           remove = FALSE) %>% 
  # remove preceding \
  mutate(company_name = 
           str_remove_all(string = company_name, pattern = "^/")) %>% 
  # check our work
  select(company, company_name, company_junk)
```

This gives us only the name in `company_name`. We can assign and remove `company_junk`.

```{r assign-company_name}
TopPharmComp %>% 
  # separate company into 2 columns
  separate(col = company, into = c("company_name", "company_junk"), 
           sep = "\\[",
           remove = FALSE) %>% 
  # remove preceding \
  mutate(company_name = 
           str_remove_all(string = company_name, pattern = "^/")) %>% 
  # reorganize
  select(rank, 
         contains("comp"),
         contains("stock"),
         contains("large"),
         everything(),
         -c(company, company_junk)) -> TopPharmComp
TopPharmComp
```

### `ranking`

Here we can use `str_remove_all()` to remove the extra numbers in `rank` and create the `ranking` variable.

```{r check-ranking}
TopPharmComp %>% 
  dplyr::mutate(ranking = 
          stringr::str_remove_all(string = rank, pattern = " \\d$")) %>% 
  select(contains("rank"))
```

We assign this to `TopPharmComp` and move onto the missing values.

```{r assign-ranking}
TopPharmComp <- TopPharmComp %>% 
  dplyr::mutate(ranking = 
          stringr::str_remove_all(string = rank, pattern = " \\d$")) %>% 
  select(ranking, 
         contains("comp"),
         contains("stock"),
         contains("large"),
         everything(),
         # remove rank
         -rank)
TopPharmComp
```

### Missing Values

We can see there is a missing value in `market_cap_us_bil` formatted as `-`. We can change these with `dplyr::na_if()`

```{r check-na_if}
TopPharmComp %>% 
  na_if(x = ., y = "-") %>% 
  select(market_cap_us_bil)
```

We can assign this and move onto formatting the variables. 

```{r assign-na_if}
TopPharmComp <- TopPharmComp %>% 
  na_if(x = ., y = "-") 
```

## Format Values {.tabset}

Let's take a look with `glimpse()` to see how these are formatted. 

```{r TopPharmComp-glimpse}
# set width
options(width = 60)
# view transposed data
TopPharmComp %>% glimpse(60)
```

We can see the `market_cap_us_bil` and `largest_market_cap_us_bil` are formatted as `character`s (but they should be `double`).

### Remove whitespace  

Remove the whitespace from `largest_market_cap_us_bil` with `stringr::str_trim()`.

```{r check-str_trim}
TopPharmComp %>% 
  mutate(
    largest_market_cap_us_bil = str_trim(string = largest_market_cap_us_bil, 
                                         side = "both"),
    largest_market_cap_us_bil = as.numeric(largest_market_cap_us_bil))
```

This is working on the `largest_market_cap_us_bil`. We will also format `market_cap_us_bil` as a numeric value.

```{r check-str_trim-as-numeric-check-market_cap_us_bil}
TopPharmComp %>% 
  mutate(
    largest_market_cap_us_bil = str_trim(string = largest_market_cap_us_bil, 
                                         side = "both"),
    largest_market_cap_us_bil = as.numeric(largest_market_cap_us_bil), 
    market_cap_us_bil = as.numeric(market_cap_us_bil))
```

Assign this to `TopPharmComp`.

```{r assign-str_trim}
TopPharmComp <- TopPharmComp %>% 
  mutate(
    largest_market_cap_us_bil = str_trim(string = largest_market_cap_us_bil, 
                                         side = "both"),
    largest_market_cap_us_bil = as.numeric(largest_market_cap_us_bil), 
    market_cap_us_bil = as.numeric(market_cap_us_bil))
```


### Dates  

We will use the `_original` suffix again to check our work. 

```{r check-lubridate-my}
TopPharmComp %>% 
  rename(cap_date_original = largest_market_cap_date) %>% 
  mutate(largest_market_cap_date = lubridate::my(cap_date_original)) %>% 
  dplyr::select(contains("date"))
```

This is giving us the date in `largest_market_cap_date`, but each day is rounded to the first day of the relative month (which is OK). Assign to `TopPharmComp` (but keep `cap_date_original`).

```{r assign-lubridate-my}
TopPharmComp <- TopPharmComp %>% 
  # rename
  rename(cap_date_original = largest_market_cap_date) %>% 
  # change format 
  mutate(largest_market_cap_date = lubridate::my(cap_date_original)) %>% 
  # reorganize
  select(ranking, 
         contains("comp"),
         contains("stock"),
         contains("date"),
         year,
         contains("us_bil"))
TopPharmComp
```

### Factors  

We want to keep the order of `ranking`, so we will format this as an ordered factor. We don't need to supply any `levels` because they are already numeric. 

```{r factor-ranking}
TopPharmComp <- TopPharmComp %>% 
  mutate(
    ranking = as.integer(ranking),
    ranking = factor(ranking, ordered = TRUE))
# check
TopPharmComp %>% arrange(ranking)
```

# Export data 

We now have a dataset we can use for modeling or visualizations! We'll export this to the `data` folder and use it in the data visualization section. 

```{r export-TopPharmComp}
readr::write_csv(x = TopPharmComp, file = "data/TopPharmComp.csv")
```

