---
title: "Exercises"
---

```{r setup, include = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
library(htmltools)
tagList(rmarkdown::html_dependency_font_awesome())
```

```{r file-list, include = FALSE, warning = FALSE, eval=FALSE}
# first this defines three functions (has_all_things, parse_yaml, ext_file,  
# and cat_bullet)
has_all_things <- function(y, things) {
  all(match(tolower(things), tolower(y), nomatch = 0) > 0)
}

parse_yaml <- function(file, n = 30) {
  rmarkdown:::parse_yaml_front_matter(read_lines(file, n_max = n))
}

ext_file <- function(file, ext) {
  paste(file, ext, sep = ".")
}
```

```{r create-basenames_extensions, warning=FALSE, include=FALSE, eval=FALSE}
basenames_extensions <- tibble(fls = list.files("exercises", 
                                                recursive = TRUE)) %>% 
  mutate(basename = tools::file_path_sans_ext(fls),
         basename = paste0("exercises/", basename),
         ext = tools::file_ext(fls)) %>% 
  group_by(basename) %>%
  summarize_all(list) %>%
  filter(map_lgl(ext, has_all_things, things = c("Rmd", "html")))
```

```{r create-hrefs_bullets, warning=FALSE, include=FALSE, eval=FALSE}
hrefs_bullets <- basenames_extensions %>%
  mutate(
    yaml = basename %>% ext_file("Rmd") %>% map(parse_yaml), 
    comment = yaml %>% map_chr("comment"),
    # comment = yaml %>% map_chr("comment", .null = ""),
    comment = str_remove_all(string = comment, pattern = "Journalism 303: "),
    comment = paste0("_*", comment, "*_"),
    comment = tolower(comment),
    href = map2(ext_file(basename, "html"), comment, ~ a(href = .x, .y)) %>%
      map_chr(as.character),
    bullet = map2_chr(href, comment, ~ paste("*", .x, .y, "\n")))
# hrefs_bullets
```

```{r cat_bullet, warning=FALSE, include=FALSE, eval=FALSE}
cat_bullet <- function(x, pattern) {
  x <- hrefs_bullets %>% 
    filter(str_detect(basename, pattern = pattern)) %>%
  unnest(ext) %>%
    filter(ext == "html")
  cat(x$bullet)
}
# cat_bullet(x, "exercises/01-")
```


### Week 1 

This is an introduction to atomic vectors (the basic building block) in R: 

* <a href="https://mjfrigaard.shinyapps.io/csuc-dj-01_atomic-vectors/">_**an introduction to atomic vectors in R**_</a> _**basic data structures**_ 

S3 vectors are slightly more complex (these include factors, dates, date-times, and difftimes), have multiple classes and attributes, and are an essential part of data analysis and visualization:

* <a href="https://mjfrigaard.shinyapps.io/csuc-dj-02_s3-atomic-vectors/">_**Introduction to S3 atomic vectors**_</a> _**factors, dates, date-times and difftimes**_ 

These exercises introduce `data.frame`s and `tibble`s, rectangular data objects in R:

* <a href="https://mjfrigaard.shinyapps.io/csuc-dj-02_data-frames/">_**rectangular data in R**_</a> _**data.frames and tibbles in R**_ 

### Week 2 

After loading or creating a data object in R, we need to be able to extract or manipulate these data. These exercises cover subsetting in R:

* <a href="https://mjfrigaard.shinyapps.io/csuc-dj-03_subsetting/">_**subsetting data in R**_</a> _**extracting data from objects in R**_ 

Nearly all of the data we'll encounter in R will be rectangular (like a spreadsheet). In R, the two forms of rectangular are `data.frame`s and `tibble`s.

* <a href="https://mjfrigaard.shinyapps.io/csuc-dj-04_rectangular-data/">_**rectangular data in R**_</a> _**learn about data.frames and tibbles, two forms of rectangular data in R**_ 

### Week 3 

`ggplot2` is an excellent package for building data visualizations and graphs in R. These exercises introduce you to the grammar and syntax. 

* <a href="">_**subsetting data in R**_</a> _**extracting data from objects in R**_ 

### Week 4 

### Week 5 

### Week 6 

### Week 7 

### Week 8 

### Week 9 

### Week 10 

### Week 11 

### Week 12 

### Week 13 

### Week 14 

### Week 15 

### Week 16 




This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](https://creativecommons.org/licenses/by-sa/4.0/).

