---
title: "Lessons"
---

```{r setup, include = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
library(htmltools)
tagList(rmarkdown::html_dependency_font_awesome())
```

```{r file-list, include = FALSE, warning = FALSE}
# first this defines three functions (has_all_things, parse_yaml, ext_file,  
# and cat_bullet)
has_all_things <- function(y, things) {
  all(match(tolower(things), tolower(y), nomatch = 0) > 0)
}

parse_yaml <- function(file, n = 30) {
  rmarkdown:::parse_yaml_front_matter(read_lines(file, n_max = n))
}

ext_file <- function(file, ext) {
  paste(file, ext, sep = ".")
}
```

```{r create-basenames_extensions, warning=FALSE, include=FALSE}
basenames_extensions <- tibble(fls = list.files("lessons")) %>%
  mutate(basename = tools::file_path_sans_ext(fls),
         basename = paste0("lessons/", basename),
         ext = tools::file_ext(fls)) %>% 
  group_by(basename) %>% 
  summarize_all(list) %>% 
  filter(map_lgl(ext, has_all_things, things = c("Rmd", "html")))

hrefs_bullets <- basenames_extensions %>% 
  mutate(
    yaml = basename %>% ext_file("Rmd") %>% map(parse_yaml),
    title = yaml %>% map_chr("title"),
    comment = yaml %>% map_chr("comment", .null = ""),
    comment = str_remove_all(string = comment, pattern = "Journalism 303: "),
    # comment = paste0("_*", comment, "*_"),
    comment = tolower(comment),
    href = map2(ext_file(basename, "html"), title, ~ a(href = .x, .y)) %>%
      map_chr(as.character),
    bullet = map2_chr(href, comment, ~ paste("*", .x, .y, "\n")))
# hrefs_bullets
```

```{r cat_bullet, warning=FALSE, include=FALSE}
cat_bullet <- function(x, pattern) {
  x <- hrefs_bullets %>% 
    filter(str_detect(basename, pattern = pattern)) %>%
  unnest(ext) %>%
    filter(ext == "html")
  cat(x$bullet)
}
```

### Week 1 

***Under construction***

Welcome! These first lessons will help you differentiate between different file types, absolute vs. relative folder paths, and installing R and RStudio:

```{r wk1-files, results='asis', echo=FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk1-")
```

### Week 2 

This week we will cover the RStudio IDE. Most of your work in the course will be done in this environment, so it's important to know to navigate the different panes and features. 

```{r wk2-files, results='asis', echo=FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk2-")
```

### Week 3 

***Under construction***

I've written a small R package, [`goodenuffR`](https://github.com/mjfrigaard/goodenuffR), to help organize your code, data, and outputs. It's based off of the excellent paper, [Good enough practices in scientific computing](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510), 

```{r wk3-files, results='asis', echo=FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk3-")
```

### Week 4 

Now that we've covered file organization and the RStudio IDE, we can move onto some R programming. These exercises cover basic R structures (vectors and functions), and help you understand how to extract pieces of your data in R. 

```{r wk4-files, results = 'asis', echo = FALSE}
cat_bullet(x, "^lessons/wk4-")
```

### Week 5 

These lessons cover how to use the R Markdown, a "*notebook interface to weave together narrative text and code to produce elegantly formatted output.*" 

```{r wk5-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk5-")
```

### Week 6 

So far, we've covered some basic R syntax. You've seen common R data objects (vectors, matrices, arrays, data frames, tibbles, and lists) and functions (`str()`, `typeof()`, `class()`, etc.). 

The `tidyverse` is an "*opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.*" Ninety percent of the code you'll see in this course uses the `tidyverse` philosophy, which makes it easier to write (and read). 

```{r wk6-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk6-")
```

### Week 7 

Now that we've introduced the tidyverse (and the `%>%`), we're going to extend this knowledge to `ggplot2`, a popular data visualization package. 

```{r wk7-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk7-")
```

### Week 8 

Data are rarely collected in a way that's fit for analysis or data visualization. In this lesson we introduce [`dplyr`](https://dplyr.tidyverse.org/), a package designed to handle "*most common data manipulation challenges.*"

```{r wk8-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk8-")
```

### Week 9 

This lesson covers intermediate data visualization techniques like plotting trends ([`geom_line()`](https://ggplot2.tidyverse.org/reference/geom_path.html)), adding help-text to your graph ([`ggplot2::annotate()`](https://ggplot2.tidyverse.org/reference/annotate.html)), labeling values to stand out ([`ggrepel::geom_label_repel()`](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html)), and using reference lines for comparisons ([`ggplot2::geom_hline()`](https://ggplot2.tidyverse.org/reference/geom_abline.html)).

```{r wk9-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk9-")
```

### Week 10 

`tidyr` has functions for *generate sequences of data* ([`full_seq()`](https://tidyr.tidyverse.org/reference/full_seq.html)), *creating an output for each combination of any input* ([`expand_grid()`](https://tidyr.tidyverse.org/reference/expand_grid.html)), and *completing missing combinations of data* ([`complete()`](https://tidyr.tidyverse.org/reference/complete.html)). 

```{r wk10-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk10-")
```

### Week 11 

This lesson builds on our previous knowledge of `ggplot2` and introduces facets (small multiples) with [`ggplot2::facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html), [`ggforce::facet_wrap_paginate()`](https://ggforce.data-imaginist.com/reference/facet_wrap_paginate.html), and [`geofacet::facet_geo()`](https://hafen.github.io/geofacet/reference/facet_geo.html)

```{r wk11-files, results = 'asis', echo = FALSE, eval=TRUE}
# and the `pr` for programming
cat_bullet(x, "^lessons/wk11-")
```

### Week 12 

Now that we've covered how to manipulate and restructure data, we'll use R's functions for creating sequences (`seq()`), along with some of `tidyr`'s functions for dealing with missing or incomplete data ([`full_seq()`](https://tidyr.tidyverse.org/reference/full_seq.html), [`expand_grid()`](https://tidyr.tidyverse.org/reference/expand_grid.html), and [`complete()`](https://tidyr.tidyverse.org/reference/complete.html)).

```{r wk12-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk12-")
```

### Week 13 

Maps are a great way to give context to any data source. In this lesson, we're going to collect data with the [`rtweet` package](https://github.com/ropensci/rtweet) and use it to build a map with [`ggplot2`](https://ggplot2.tidyverse.org/reference/geom_map.html).

```{r wk13-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk13-")
```

### Week 14 

***Under construction***

Not every project requires a data visualization, so this week we'll cover how to display data in tables. 

```{r wk14-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/wk14-")
```

### Week 15 

```{r cs01-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/cs01-")
```

### Week 16 

```{r cs02-files, results = 'asis', echo = FALSE, eval=TRUE}
cat_bullet(x, "^lessons/cs02-")
```




This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](https://creativecommons.org/licenses/by-sa/4.0/).

<!--





This week we cover our first case study using data from the executed offenders database on the [Texas Department of Criminal Justice website](https://www.tdcj.texas.gov/death_row/dr_executed_offenders.html).










> Now that we've introduced the grammar of `ggplot2`, we can build visualizations that display counts and variable distributions. 

> In previous lessons, we used `dplyr` to manipulate data. In this lesson, we'll be using the `tidyr` package to reshape and restructure data. To demonstrate how these packages work together, we're going to scrape (download) a Wikipedia table of [List of largest biotechnology and pharmaceutical companies](https://en.wikipedia.org/wiki/List_of_largest_biotechnology_and_pharmaceutical_companies). 
--> 
